---
title: "External Source Attribution and Adaptation while Viewing the Hand"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r load_code, include=FALSE}
source('R/shared.R')
source('R/learningCurves.R')
source('R/noCursor.R')
source('R/localization.R')
source('R/distributionplots.R')
source('R/awareness.R')
```


# Overview

This document discusses figures and statistics for the Hand View project. The five main sections include data and analyses with regards to Learning Curves, Reach Aftereffects, Localization, Contributions of proprioception and predictions to reach aftereffects, and Extra analyses conducted. 

# Learning Curves

First, we aimed to determine the effects of externally attributing the source of errors on reaching movements while experiencing a 30° CW rotation. 

Training reaches in *Cursor Trials* for both aligned and rotated sessions were manually inspected for failing to perform the required reach, e.g. not reaching the target with a smooth and straight movement. Trials that were not useable for further analyses were removed. For the remaining trials, we calculated angular deviation of the hand at the point of maximum velocity. This is the angular difference between a line connecting the home position to the target and a line connecting the home position to the participant's hand position at the peak of their movement velocity. Thus, in order to fully compensate for the cursor-rotation, participants had to deviate their reaches in the 30° CCW direction. We then calculated the difference between angular deviations in the first 90 trials of the rotated session and individual biases in reaches during the last 30 out of the first 45 trials in the aligned session. This process was completed for every participant in each group.


## Show the data

Below is a plot of the angular deviation of the hand across the first 90 trials of the rotated session. Lines represent the mean angular deviation of each group in every trial, and corresponding shaded regions represent 95% Confidence Intervals. For a closer look, we also show the same data, but only for the first 15 and last 15 trials.

```{r plot_learningCurves}
plotLearningCurvesOLD(target='inline')
plotBlockedLearningCurves(target='inline')
```

We see that, during training, the Instructed Group countered the rotation immediately, while the other groups took longer (within 15-20 trials) to compensate for the cursor-rotation. The Cursor Jump group's rate of learning seems to suggest that their rate of learning was much faster than the Non-instructed, control group (i.e. the slope of the learning curve is steep). Lastly, all groups performed similar reaching movements at the end of training.

## ANOVAs

To analyze the effects of our manipulation on rates of learning, we blocked the rotated training data into three blocks of trial sets. The first block consisted of the first set of three trials, the second consisted of the next three trials, and the third consisted of the last 15 trials of rotated training. Below is a plot showing individual participant data for each group across these three blocks.

```{r}
plotLearningCurves()
```

From the plot, we see that the distribution of the angular hand deviation across all participants are relatively normal and unimodal for all groups and blocks. Therefore, we proceeded to investigate how the four groups and their performance during adaptation training, compared with each other, for each of these three blocks. We performed a 3X4 mixed design ANOVA with block as a within-subject factor (blocks 1, 2, and 3) and group as a between-subject factor (Non-instructed, Instructed, Cursor Jump, Hand View). For this and the following tests, the alpha level was set to 0.05 and, when necessary, Greenhouse Geisser corrections were applied.

```{r}
learningCurveANOVA()
```
We see that the main effects of group and block, and their interaction, were statistically significant. This suggests that the rate of learning of at least one group differed from the other groups and that this difference was dependent on the block of trials. Hence, it was appropriate to look at differences between groups in each block. Particularly, we only looked into the first 2 blocks, since it was observed from our plots that the groups did not differ in the last block.

## Planned follow-up tests

We first display the means and 95% CI's for each group in each block. This table gives us an idea of which contrasts we want to test further. That is, if the mean of a group falls outside the confidence interval range of another group then the two groups must differ significantly from each other. For these contrasts, we use the Non-instructed group as a reference (since it is the control group) and compare all the other groups to it. We then compare all groups to the non-instructed group for blocks one and two, respectively. To control for multiplicity, we adjusted for p values using the Sidak method. Note that the Sidak method is the default used in R when an alternative test, such as Tukey, is inappropriate for a given dataset. The Tukey method would be appropriate if we were comparing each mean with every other mean, but comparing only a selection of means would require the Bonferroni or Sidak. The Sidak method offers the slight advantage of being a less conservative method than the Bonferroni. Another option would be to use the Dunnett method when comparing a group with a control group. Nevertheless, all of these are just for adjusting p values. Our results do not change, however, regardless of the method used in terms of which comparisons reached significance. Hence, we chose to stick with the default Sidak method.

```{r}
learningcurveComparisonMeans()
```

```{r}
getComparisonEffSize(method='sidak')
```

In block one, when each group was compared to the Non-instructed group, only the Instructed group differed significantly. This suggests that only the instructed group had an initial advantage in learning to counter for the rotation of the cursor. In block two, however, only the cursor jump group significantly differed from the non-instructed group, suggesting that the cursor jump group learned to counter for the rotation of the cursor at a faster rate than the non-instructed group. The hand view group, however, did not differ significantly from the non-instructed group. Note that effect sizes are provided with eta-squared, which is interpreted as the percent of variance in angular deviation of the hand, accounted for by the difference between the given group and the control, Non-Instructed group.

To summarize, instruction on the nature of the perturbation and being given a strategy to counter it, creates an initial advantage in reducing hand direction error when reaching towards a target with a rotated cursor. However, our other manipulations, despite the error being more clearly external in nature, do not produce a similar initial advantage.


# Reach aftereffects

We then investigated how manipulating the extent of external error attribution affected participants' reaching movements when no cursor-feedback is presented.

Similar to the Learning Curves section, reaches in *No Cursor Trials* for both aligned and rotated sessions were manually inspected for failing to perform the required reach. Trials that were not useable for further analyses were removed. For the retained trials, we used the endpoint angle to calculate for the angular deviation of the hand. This is the angular difference between the point where the participant's hand movement ended and the target. We then subtracted mean angular deviations in the aligned session from those after training in the rotated session, in order to measure the effects of adaptation training during No Cursor reaches, i.e. reach aftereffects. We then implemented a process dissociation procedure (PDP) type of analysis. We calculated mean angular deviations per participant when either including any strategy they learned during adaptation training or excluding such a strategy. Thus, awareness of learning about the visuomotor rotation would be associated with a difference between exclusion and inclusion trials, while lack of awareness will have no such difference between the two types of trials.

## Show the data

Below is a plot of the mean angular reach deviation of the hand for each group as a function of either excluding or including a strategy. Lines represent the mean angular deviation of each group, and corresponding shaded regions represent 95% Confidence Intervals.

```{r plot_noCursor}
plotGroupReachAfterEffects(target='inline')
```

We see from the figure that when including or excluding the strategy learned to counter the rotation, only the non-instructed group could not switch their strategy on or off at will. This suggests unawareness of learning for this group. Importantly, all groups still showed angular deviations of their hand to some degree during exclusive trials. This suggests that despite awareness of learning of the visuomotor rotation, implicit reach aftereffects were still observed in all groups. Note, however, that implicit reach aftereffects were smaller for the Hand View group compared to the other groups.

Below is a plot showing individual data points for reach aftereffects and strategy use across all groups.

```{r}
plotReachAfterEffects()
```

We see that distributions for all groups are normal and unimodal. Thus, we proceeded with the statistical tests performed below.

## Test for Reach Aftereffects

First, we tested whether or not adaptation led to reach aftereffects. To do this, we compared the rotated no cursor trials to the aligned no cursor trials. Particularly, we used rotated no cursor trials where participants did not make use of a strategy. We performed a 2X4 mixed design ANOVA with session (aligned or rotated) as a within-subject factor and group as a between-subject factor.

```{r}
NoCANOVA()
```

We see that there were implicit reach aftereffects after adaptation due to the significant main effect of session. However, since the interaction between group and session was also significant, it is possible that the effect of session was modulated by groups (i.e. the effect of session occurs differently for at least one group). Thus, we performed planned follow-up tests comparing each group to itself between the two sessions.

```{r}
NoCursorComparisonMeans()
getNoCurComparisonEffSize(method='sidak')
```

Planned follow-up tests show that the aligned and rotated sessions differed significantly for each group. Note that this test does not definitively tell us which of the groups differed from the others, in terms of differences between sessions. However, we see from our figure that it is the Hand View group which exhibited a lower angular deviation of the hand in exclusive trials compared to the other three groups. For the purpose of answering the question we are investigating in this subsection, it is sufficient that the follow-ups showed that sessions differed for each group. Note that effect sizes are provided with eta-squared, which is interpreted as the percent of variance in angular deviation of the hand, accounted for by the difference between the two sessions.

## PDP-type analysis (Strategy Use)

After confirming that reach aftereffects were present in all groups, we tested whether our manipulation of the extent of external error attribution differed across groups, when either excluding or including a strategy to counter for the rotation. We performed a 2X4 mixed design ANOVA with strategy use as a within-subject factor (with or without strategy) and group as a between-subject factor.

```{r}
RAEANOVA()
```

We see that the main effects of group and strategy, as well as their interaction were significant. This suggests that angular reach deviations from at least one group differed from the other groups, when either using or not using a strategy. However, planned follow-up tests were necessary to determine which of the groups differed.

```{r}
RAEComparisonMeans()
getRAEComparisonEffSize(method='sidak')
```

The tests confirmed our conclusions from the figure. All groups, except the non-instructed group, showed significantly different angular reach deviations when comparing reaches using a strategy to reaches when they are not using a strategy. This suggests unawareness of learning for the non-instructed group. Note that effect sizes are provided with eta-squared, which is interpreted as the percent of variance in angular deviation of the hand, accounted for by the difference between using and not using a strategy.

# Localization

We also investigated how manipulating the extent of external error attribution affected participants' hand position estimates for both before and after adaptation training.

we analyzed estimates of hand position in *Active Localization* and *Passive Localization* trials, before and after rotation training. In obtaining hand position estimates, we calculated the angular difference between the point where the participant’s unseen right hand movement ended and their perceived hand position, i.e. their touchscreen response using the left hand. For us to be able to calculate these angular difference measures appropriately, the touchscreen and monitor position must be coordinated. However, during our analyses of the data, we encountered inconsistencies with the location of the touchscreen responses relative to the arc location. This is because the touchscreen position is fixed, but the monitor can be moved to different positions. Thus, these inconsistencies showed that some of the responses were not centred on where the arc stimuli were expected to be displayed (12 cm from home position). Despite this, responses still fell within the shape of an arc, which allowed us to fix for this problem by implementing a circle fitting procedure that finds a circle of best fit for the responses. Moreover, this procedure also allowed us to account for any participant response biases. The procedure used a fitting algorithm that minimized mean squared errors (MSE), where errors were defined as the difference between the touchscreen response locations and the expected arc stimulus location. The MSE measures were then passed on to an optimizing function (optim in R), which minimized the error to return a best fitting home position, such that the responses were shifted to where we expected it to be (i.e. 12 cm from home position). In short, the circle fitting procedure ensured that any localization shifts we detected in our analyses were not due to technical issues or response biases. Furthermore, we conducted an outlier removal procedure for this fitted data, such that touchscreen responses beyond ±20° from the arc centre and angular errors beyond ±3 standard deviations from the mean angular error were removed. Lastly, as participants chose the points on the arc that they moved towards, their movements did not end on all possible points on the arc. To account for this, we used a kernel smoothing method (width = 15°) to interpolate changes in hand localization at specific interpoints (50°, 90°, and 130°). Note that these interpoints are the same points where the arc stimuli are centred on in polar coordinates. We then calculated for the mean of these interpolated values, at every interpoint, to estimate shifts in hand localization.


## Show the data

Below is a plot of the mean angular shift in hand localization for each group as a function of the three interpoints (50°, 90°, and 130°). Panel 1 shows shifts in hand localization when both proprioceptive information and efference-based predictions are available to the participant. Panel 2 shows shifts in hand localization when only proprioceptive information is available. Panel 3 shows shifts in hand localization when efference-based predictions are isolated (i.e., shifts in Active minus shifts in Passive Localization). Lines represent the mean angular shift in localization for each group, and corresponding shaded regions represent 95% Confidence Intervals. The right hand side of each panel shows the mean localization shifts and corresponding 95% Confidence Intervals for each group, regardless of Hand Angle.

```{r plot_localization}
plotLocalizationShift(target='inline')
```

We see from the figure that all groups showed a shift in hand localization, and that these shifts were larger when both efference-based predicitions and proprioceptive information were available to the participants (i.e., *Active Localization*). In comparing Panels 2 and 3, we see that majority of the shifts in active localization was accounted for by proprioceptive information, as the shifts in predicted sensory consequences were near 0. However, the lack of differences across groups seems to suggest that our manipulation of external error attribution did not affect both proprioceptive recalibration and updating of predicted sensory consequences. Before any statistical tests, however, we first checked for the shape of distributions of individual data points across groups. Below is a plot showing individual data points for shifts in localization according to groups and movement type.

```{r}
plotLocalizations()
```

We see that distributions for all groups are normal and unimodal. Thus, we proceeded with the statistical tests performed below.

## Localization shifts after rotated training

First, we wanted to test whether localization estimates did shift after training with the rotated cursor. We ran a 2X2X4 mixed design ANOVA with session (*rotated_b* in data; aligned and rotated) and movement type (*passive_b* in data; active and passive) as within-subjects factors and group (instructed, non-instructed, cursor jump, and hand view) as a between-subject factor.

```{r}
localizationANOVA(test='omnibus')
```

To answer the question for this sub-section, we see that rotated_b had a significant main effect, which suggests that estimates of localization after rotated training differed from localization estimates before training. We also see other effects which we will investigate further.

## Effect of external attribution and movement type on localization shifts

Since localization estimates shifted after rotated training, we tested whether these shifts were affected by our different groups (i.e. our manipulation of the extent of external error attribution) and by the movement type executed (i.e. active or passive). We ran a 2X4 mixed design ANOVA with movement type (*passive_b*) as a within-subject factor and groups as a between-subject factor. 

```{r}
localizationANOVA(test = 'shifts')
```

We see that there is a main effect of group and movement type, but no significant interaction. This suggests that the effects of movement type on shifts in hand position estimates were not modulated by group, and vice-versa. Our main research question aimed to see changes in estimates of hand location due to proprioceptive recalibration and updating of predicted sensory consequences, which we investigated next.

## Proprioceptive Recalibration

As previously stated, during active localization, participants have access to both information on proprioception and predicted sensory consequences. On the other hand, during passive localization, participants only have access to proprioceptive information. We are interested in testing the effect of manipulating the extent of external error attribution (group) on both proprioceptive recalibration and updates in predicted sensory  consequences.

First, we run a one way ANOVA to test for the effect of group on passive localization changes (proprioceptive recalibration).

```{r}
localizationANOVA(test='passive')
```

We see that there are no significant differences across the groups. This suggests that passive localization changes were not affected by our manipulation of error attribution. *NOTE: (In the instance that we might be interested in testing for the effect of group on active localization changes, the function can also run the one way ANOVA on active localization changes. I do not show it here, but in that test, there is a significant difference between groups).*

Next, we wanted to test whether each group showed any recalibration in proprioception. We ran t-tests that compared each group's magnitude of shifts in passive localization estimates against 0. These were one-tailed t-tests, where the null hypothesis was that the mean is greater than or equal to 0; the alternative hypothesis was that the mean is less than 0. Hence confidence intervals are all from minus infinity. Note that we do not need to control for multiplicity here, as each group is independently compared to zero (i.e., these are not pairwise comparisons).

```{r}
pasLocTtests()
```

From the tests, we see that all groups differed significantly from 0, suggesting that all groups showed changes in proprioceptive estimates. Note that effect sizes are provided with Cohen's d, which is in standard deviation units. Thus, it indicates how many standard deviations the mean shift for a group is different from zero. Here, all Cohen's d values are relatively similar, except for the Hand View group.


## Predicted Sensory Consequences

Next, we ran a one way ANOVA to test for the effect of group on changes in predicted sensory consequences. For this ANOVA, we obtained changes in predicted sensory consequences by subtracting the shifts in passive localization from active localization.

```{r}
predictedConsequencesANOVA()
```

There were no significant differences across the groups, suggesting that changes in predicted sensory consequences were not affected by our manipulation of error attribution.

However, from the plot we generated above, it seems that only the Hand View group is at the 0-degree mark for changes in predicted sensory consequences. Thus, we tested whether each group showed any predicted sensory consequences at all, especially for the Hand View group. We ran t-tests that compared each group's magnitude of changes in predicted sensory consequences against 0.

```{r}
predConsTtests()
```

From the tests, we see that only the Hand View group did not significantly differ from 0, suggesting that the Hand View group did not show any updates in predicted sensory consequences. Note that effect sizes are provided with Cohen's d, which is in standard deviation units. Thus, it indicates how many standard deviations the mean shift for a group is different from zero. Here, only the Hand View group has a very small effect size.

# Contributions of Proprioception and Predicted Sensory Consequences on Reach Aftereffects

According to Modchalingam et al. (2019), afferent and efferent-based changes in hand localization are independent processes, and they suggested that afferent-based changes (proprioceptive recalibration) partly predicts implicit motor changes (reach aftereffects). Here, we test whether both afferent and efferent-based changes predict/contribute to implicit motor changes. 

To visualize the data, we plot the relationship of proprioception and aftereffects, as well as predictions and aftereffects, individually. Note that we did perform a Pearson product moment correlation test for these two and found a significant relationship for both, but we only include it as analyses extra to the manuscript (i.e., the main test is the multiple regression below).
```{r}
plotPropPredRelationships()
```

From the left and centre panel of the plot, we see that there seems to be a relationship between predictions and aftereffects, as well as proprioception and aftereffects. For these plots, we pooled data from all groups together. One criticism could be that we are observing a spurious correlation due to all groups being included. However, as we are investigating adaptation and its underlying processes, we would want to see any differences in groups with regards to how they are learning. Therefore, we pooled the data across all groups to get a sense of whether there is a continuum between groups that show these differences in learning. If we considered the groups individually, the range of adaptation would be narrow because they are doing the same tasks, and any potential effects seen here would mostly be due to noise. we note that there are no prior differences between groups, as participants were randomly sampled, all did the same task, and experienced a similar perturbation magnitude. Therefore, any differences in groups here are due to the manipulation in the experiment. (See extra analyses below).

We then performed a multiple regression, where measures of aftereffects was the dependent variable, and both proprioceptive recalibration and predicted sensory consequences were the predictors.

```{r}
getPropPredGLM()
```

We found that both changes in proprioception (p < .05) and predictions (p < .05) were significant predictors of reach aftereffects. Importantly, we note that changes in proprioception are independent from predicted sensory consequences due to how we have measured them. Recall that proprioceptive recalibration was measured from Passive localization, while predictions were calculated as a difference of Active and Passive Localization. This process ensures predictions do not have any shared variance with proprioceptive recalibration. Furthermore, predictions and proprioceptive changes are not collinear with each other (vif = 1.09 is considered low; see extra analyses below). 

Next, we calculated the predicted reach aftereffects from our model and compared these with the actual/observed reach aftereffects from our experiment (right panel in the plot above).To calculate the predicted reach aftereffects, we refer to values from our model. The equation is:

predicted RAE = Intercept + Prediction_Coeff(actual_prediction) + Proprioception_Coeff(actual_proprioception)

If the model predicted reach aftereffects perfectly, then points would fall along the diagonal of the plot. We see that the predictions overshoot the data on the lower end, but undershoot the data on the higher end. These could possibly be due to noise, or it could mean that the model is incomplete. It is very likely that another predictor needs to be included in the model, but the current study did not look further into this.

# Analyses extra to those included in the Manuscript

## Awareness Scores

### Show scores for all groups

Below are two plots showing the data on awareness scores by group. The barplot shows the number of participants that scored either 0, 1, 2, or 3 on the Awareness Questionnaire, for each of the four groups. The dot plot generated beside it, gives a visual of the number of participants scoring either high or low on the questionnaire, separated according to group.

```{r}
plotAwareness()
```

The plots confirm our findings with the PDP-type analysis discussed above. The Non-Instructed group scored lower on the questionnaire compared to the other groups, suggesting unawareness of the nature of the perturbation.

## Pearson correlation test between implicit reach aftereffects and passive localization

Here, we tested whether the shifts in passive localization are correlated with the angular reach deviations of the hand in *No Cursor - Without Strategy* trials. We performed a Pearson product-moment correlation test, which included all particiants regardless of group.

```{r}
getRAEPropCorrelation()
```

We found a negative relationship (r = -0.35) between afferent-based changes in localization and reach aftereffects when excluding a strategy (p < 0.001, r-squared = 0.12). This relationship suggests that proprioceptive recalibration could be contributing to the reach aftereffects we have observed.

We performed the same test for updates in predicted sensory consequences and reach aftereffects. 

```{r}
getRAEPredCorrelation()
```

We found a negative realtionship (r = -0.30) between efferent-based changes in localization and reach aftereffects when excluding a strategy (p < 0.01, r-squared = 0.09). This relationship suggests that updates in predictions are also contributing to reach aftereffects. We suggest that the contribution of predictions are independent from the contributions of proprioceptive recalibration.

## Additional analysis for correlations/regression

To show that changes in predictions and proprioception are independent predictors/contributors to reach aftereffects, proprioception should be able to predict the residuals of regressing aftereffects on predictions.

```{r}
getPredResidByPropCorrelation()
```

We find a significant relationship between proprioception and residuals of regressing aftereffects on predictions (p < 0.001, r = -0.45). We can also see if the same is true for predictions and residuals of regressing aftereffects on proprioception. 

```{r}
getPropResidByPredCorrelation()
```

We find a significant relationship as well (p < 0.001, r = -0.42). Lastly, we can compare these two relationships with our original correlational plots.

```{r}
plotRelationships()
```

By considering the residuals, we are essentially controlling for one predictor. However, the existence of a significant relationship, between one predictor and the residuals of aftereffects with the other predictor, suggests that the two predictors are independent from each other.

## Further analysis on the low VIF scores from Multiple Regression

The low VIF score from predictions and proprioceptive changes could be trivial, given that predicted sensory consequences are measured by subtracting Passive Localization scores from Active Localization (PSC = ACT - PAS). To investigate this further, we could simulate either ACT or PAS scores that will also have a low VIF score with the other variable, but will not correlate with the residuals of regressing aftereffects on the other variable.

First, we generated simulated PAS scores. To do this, we shuffled the PSC scores then subtracted them from ACT scores. Then we tested whether these simulated PAS scores will correlate with the residuals of regressing reach aftereffects with ACT scores.

```{r}
getACTResidByFakePASCorrelation()
```

We found that the simulated PAS scores do not significantly correlate (p = 0.81, r = -0.03) with the residuals of regressing aftereffects on ACT, even if ACT was a significant predictor of aftereffects (p < 0.001). The VIF score between the simulated PAS scores and PSC was still low (vif = 1.11). We also repeated this analysis the other way around and generated simulated ACT scores instead. 

```{r}
getPASResidByFakeACTCorrelation()
```


We found that the simulated ACT scores do not significantly correlate (p = 0.92, r = 0.01) with the residuals of regressing aftereffects on PAS, even if PAS was a significant predictor of aftereffects (p < 0.001). The VIF score between the simulated ACT scores and PSC was still low (vif = 1.08). 

Another simple check is to show that ACT and PAS VIF scores are high, as these are considered to be collinear. This is exactly what we found (vif = 3.37), as shown below.

```{r}
getActPasCollinearity()
```

Given the low VIF score we found for predictions and proprioception, but high VIF score for ACT and PAS, we are certain that the VIF score informs uf about the collinearity of these variable. However, it is trivial in its ability to show the independence of predictions from proprioception. Thus, the residual analyses on the previous section above is more informative in showing that these two predictors are independently contributing to reach aftereffects. 

## Analyzing potential group effects

We also wanted to ensure that whatever effects we found from the multiple regression would be explained by group differences. Here, I display three potential ways of looking into these. 

### 1)
One way of doing this was to calculate for the mean PSC and PAS scores, then subtract these from each individual PSC and PAS score within each group. Afterwards, we can replot the data and test for the existence of a relationship between PSC and PAS with aftereffects. The absence of an effect after doing such a correction will indicate that any effects observed from the original multiple regression is explained by group differences (or the experimental manipulation).

```{r}
getMeanCorrectedGLM()
```

However, we found that both predictors (predictions: p < 0.05; proprioception: p < 0.01) were still significant after doing such a correction. We can also consider proprioception and predictions separately. First, we replot the relationship of proprioceptive recalibration with aftereffects after the correction, then test for a correlation.

```{r}
plotMeanCorrectedPropCorrelations()
getMeanCorrectedRAEPropCorrelation()
```

We found that proprioception has a significant relationship with aftereffects (p=0.03, r = -0.24), but this is less in magnitude compared to the original data (r-squared went from 0.12 to 0.06). We repeated the same analyses for predictions.

```{r}
plotMeanCorrectedPredCorrelations()
getMeanCorrectedRAEPredCorrelation()
```

We found the predictions did not have a significant relationship with aftereffects (p=0.21, r = -0.13), and r-squared values were reduced from 0.09 to 0.02.

### 2)
Another way of looking into group effects is to see how the relationships of predictions and proprioception with aftereffects change with the removal of the Hand View group (Note that we only remove the Hand View group because it was the group observed to be different from the others in measures of aftereffects and predicted sensory consequences). First, we investigate the relationship of predictions and aftereffects, without the Hand View group.

```{r}
plotNoHVPredGroupCorrelations()
getNoHVRAEPredCorrelation()
```

We found that the correlation between predictions and aftereffects, after removing the Hand View group, was not significant (p > 0.10, r = -0.21, r-squared = 0.04). This suggests that the observed effect from previous analyses could be mainly driven by the Hand View group. (Refer back to section on *Contributions of Proprioception and Predicted Sensory Consequences on Reach Aftereffects*).

Next, we repeat the analyses for proprioceptive recalibration and aftereffects.

```{r}
plotNoHVPropGroupCorrelations()
getNoHVRAEPropCorrelation()
```

We found that the correlation between proprioceptive changes and aftereffects, after removing the Hand View group, was significant (p < 0.05, r = -0.28, r-squared = 0.08).

### 3)
We also looked into the relationship of predictions and proprioception on reach aftereffects, for each individual group. First, I show results for correlation tests between proprioception and aftereffects for each group. I also pot these relationships to visualize the data better.

```{r}
getPGPropCorr()
plotPGPropCorr()
```


We see that only the Hand View group exhibits a significant relationship between proprioception and aftereffects (p < 0.01, r = -0.50), while other groups did not exhibit such a relationship (Cursor Jump: p > 0.10, r = -0.34; Instructed: p > 0.10, r = -0.29; Non-Instructed: p > 0.10, r = -0.17). 

We then repeat the same analyses for predictions and aftereffects.

```{r}
getPGPredCorr()
plotPGPredCorr()
```

We see that none of the groups exhibited a significant relationship between predictions and aftereffects (Hand View: p > 0.10, r = -0.003; Cursor Jump: p > 0.10, r = -0.009; Instructed: p > 0.10, r = -0.30; Non-Instructed: p > 0.10, r = -0.30).

Taken together, all of these analyses show that there are group differences due to the experimental manipulation, when looking into relationships between predictions and aftereffects, as well as proprioception and aftereffects. However, pooling the data together is justified due to our interest in showing what seems to be a continuum of how the groups differ in learning.
