---
title: "External Source Attribution and Adaptation while Viewing the Hand"
author: "Raphael Gastrock"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

# Overview

This document discusses figures and statistics for the Hand View project. The four main sections include data and analyses with regards to Learning Curves, Reach Aftereffects, Localization, and Contributions of proprioception and predictions to reach aftereffects.

# Setting up the R environment

Our scripts require the use of certain packages. If some packages are not available in your own system, run the command:

> install.packages(c('ez', 'svglite', 'phia', 'afex', 'lsmeans', 'scales', 'lsr', 'tools', 'usdm', 'lmtest', 'car', 'rockchalk', 'tidyr'))


We have written code that pre-processes the data, plots figures, and runs statistical tests. Import the sources below:

```{r message=FALSE, warning=FALSE}
source('R/shared.R')
source('R/learningCurves.R')
source('R/noCursor.R')
source('R/localization.R')
source('R/awareness.R')
source('R/cursortrajectory.R')
source('R/data.R')
```

Then we need to ensure that the data is available locally. We download the pre-processed data available from OSF. If the files already exist, the output will indicate this.

```{r}
getPreprocessedData()
```

Once the data and code are available, we can proceed to the next sections.

# Learning Curves

First, we aimed to determine the effects of manipulating the extent of attributing the source of errors externally, on reaching movements while experiencing a 30° CW rotation. 

Cursor training trials for both aligned and rotated sessions were manually inspected for failing to perform the required reach, e.g. not reaching the target with a smooth and straight movement. Trials that were not useable for further analyses were removed. For the remaining trials, we calculated angular deviation of the hand at the point of maximum velocity. This is the angular difference between a line connecting the home position to the target and a line connecting the home position to the participant's hand position at the peak of their movement velocity. Thus, in order to fully compensate for the cursor-rotation, participants had to deviate their reaches in the 30° CCW direction. We corrected for individual baseline biases, by calculating the average reach deviation for each target separately within each participant, during the last 30 out of the first 45 aligned cursor training trials, and subtracting this from cursor training trials in the rotated
session.


## Show the data

Below is a plot of the angular deviation of the hand across the first 90 trials of the rotated session. Lines represent the mean angular deviation of each group in every trial, and corresponding shaded regions represent 95% Confidence Intervals. For a closer look, we also show the same data, but only for the first 15 and last 15 trials.

```{r plot_learningCurves}
plotFullLearningCurves(target='inline')
plotBlockedLearningCurves(target='inline')
```

We see that, during training, the Instructed Group countered the rotation immediately, while the other groups took longer (within 15-20 trials) to compensate for the cursor-rotation. All groups performed similar reaching movements at the end of training.

We also show individual and average reach trajectories for all groups, during the aligned session, early adaptation training, and late adaptation training. From this plot, we observe that reach trajectories are similar in pattern across all groups, regardless of manipulations to the extent of external error attribution.

```{r fig.width=8, fig.height=8}
plotAllTrajectories()
```


## ANOVAs

To analyze the effects of our manipulation on rates of learning, we blocked the rotated training data into three blocks of trial sets. The first block consisted of the first set of three trials, the second consisted of the next three trials, and the third consisted of the last 15 trials of rotated training. Below is a plot showing individual participant data for each group across these three blocks (i.e. transparent dots). Solid dot and error bars represent the mean and bootstrapped 95% confidence intervals.

```{r fig.width=8, fig.height=8}
plotLearningCurves()
```

We proceeded to investigate how the four groups and their performance during adaptation training, compared with each other, for each of these three blocks. We performed a 3X4 mixed design ANOVA with block as a within-subject factor (blocks 1, 2, and 3) and group as a between-subject factor (Control, Instructed, Cursor Jump, Hand View). For this and the following tests, the alpha level was set to 0.05 and, when necessary, Greenhouse Geisser corrections were applied.

```{r}
learningCurveANOVA()
```
We see that the main effects of group and block, and their interaction, were statistically significant. This suggests that the rate of learning of at least one group differed from the other groups and that this difference was dependent on the block of trials. Hence, it was appropriate to look at differences between groups in each block.

## Planned follow-up tests

We first display the means and 95% CI's for each group in each block. This table gives us an idea of which contrasts we want to test further. That is, if the mean of a group falls outside the confidence interval range of another group then the two groups must differ significantly from each other. For these contrasts, we use the Control group as a reference and compare all the other groups to it. We then compare all groups to the Control group for blocks one and two, respectively. To control for multiplicity, we adjusted for p values using the Sidak method. Note that the Sidak method is the default used in R when an alternative test, such as Tukey, is inappropriate for a given dataset. The Tukey method would be appropriate if we were comparing each mean with every other mean, but comparing only a selection of means would require the Bonferroni or Sidak. The Sidak method offers the slight advantage of being a less conservative method than the Bonferroni. Another option would be to use the Dunnett method when comparing a group with a control group. Nevertheless, all of these are just for adjusting p values. Our results do not change, however, regardless of the method used in terms of which comparisons reached significance. Hence, we chose to stick with the default Sidak method.

```{r}
learningcurveComparisonMeans()
```

```{r}
getComparisonEffSize(method='sidak', type='two')
```

In block one, when each group was compared to the Control group, only the Instructed group differed significantly. This suggests that only the instructed group had an initial advantage in learning to counter for the rotation of the cursor. In block two, however, no groups significantly differed from the Control group, suggesting that the learning across groups did not differ from each other for this block. Note that effect sizes are provided with eta-squared, which is interpreted as the percent of variance in angular reach deviation that is accounted for by the difference between the given group and the control, Control group.

To summarize, instruction on the nature of the perturbation and being given a strategy to counter it, creates an initial advantage in reducing hand direction error when reaching towards a target with a rotated cursor. However, our other manipulations, despite the error being more clearly external in nature, do not produce a similar initial advantage.

## Confirmatory Analyses for follow-up tests

In addition to our planned follow-up tests, we also conducted extra analyses that will support our findings. First, in the multiple pairwise comparisons we conduct above, contrasts were specified specifically for blocks 1 and 2. However, if we include the last block into these contrasts and comparisons, the calculated results will differ in some manner.

```{r}
getComparisonEffSize(method='sidak', type='all')
```

The main different result would be that the Cursor Jump group is now statistically different from the Control group, within the second block. This suggests that the Cursor Jump group's rate of learning was faster than the Control group for this block, but this does not provide any further advantage, as the magnitude of asymptotic learning across all groups do not differ from each other by the end of adaptation training. Support for this comes from the lack of statistical significance for comparisons within the last block (as shown above), as well as an additional One-way ANOVA we conduct on reach deviations for only the last block (as shown below).

```{r}
lastblockANOVA()
```


# Reach aftereffects

We then investigated the effect of manipulating the extent of external error attribution on reach aftereffects.

Similar to the Learning Curves section, reaches in no-cursor trials for both aligned and rotated sessions were manually inspected for failing to perform the required reach. Trials that were not useable for further analyses were removed. For the retained trials, we used the endpoint angle to calculate for the angular deviation of the hand. This is the angular difference between the point where the participant's hand movement ended and the target. Angular reach deviations from aligned no-cursor trials were subtracted from without strategy no-cursor trials to confirm whether training with a rotated cursor led to reach aftereffects.

## Show the data

Below is a plot of the mean angular reach deviation for each group as a function of either excluding or including a strategy. Lines represent the mean angular deviation of each group, and corresponding shaded regions represent 95% Confidence Intervals. Individual participant data are shown as transparent dots.

```{r plot_noCursor}
plotGroupReachAfterEffects()
```

We see from the figure that when including or excluding the strategy learned to counter the rotation, only the Control group could not switch their strategy on or off at will. This suggests unawareness of learning for this group. Importantly, all groups still showed angular reach deviations to some degree during without strategy trials, showing that implicit reach aftereffects were still observed in all groups. Note, however, that implicit reach aftereffects were smaller for the Hand View group compared to the other groups.

Below is a plot showing individual data points for reach aftereffects and strategy use across all groups. The left panel is the same as the plot shown above. For the centre and right panels, solid dots and error bars represent the mean and bootstrapped 95% confidence intervals, and transparent dots represent individual participant data.

```{r}
plotReachAfterEffects()
```


## Test for Reach Aftereffects

First, we tested whether or not adaptation led to reach aftereffects. We performed a 2X4 mixed design ANOVA with session (aligned or rotated) as a within-subject factor and group as a between-subject factor.

```{r}
NoCANOVA()
```

We see that there were implicit reach aftereffects after adaptation due to the significant main effect of session. However, since the interaction between group and session was also significant, it is possible that the effect of session was modulated by groups (i.e. the effect of session occurs differently for at least one group). Thus, we performed planned follow-up tests comparing each group to itself between the two sessions.

```{r}
NoCursorComparisonMeans()
getNoCurComparisonEffSize(method='sidak', type='all')
```

Planned follow-up tests show that the aligned and rotated sessions differed significantly for each group. Note that effect sizes are provided with eta-squared, which is interpreted as the percent of variance in angular reach deviation that is accounted for by the difference between the two sessions.

To investigate which group differed from each other in the exclusive (without strategy) no-cursor trials, we conducted pairwise comparisons of each group compared to the Control group, as well as a comparison between the Cursor Jump and Hand View group.

```{r}
getNoCurComparisonEffSize(type='exclusive')
```

We find that the Cursor Jump and Hand View groups are significantly different from the Control group, suggesting that implicit reach aftereffects for these groups are lower. Moreover, a significant difference between the Cursor Jump and Hand View group, suggests how implicit reach aftereffects are further reduced for the Hand View group.

## PDP-type analysis (Strategy Use)

Following analyses for reach aftereffects, we assessed the effect of our manipulations of external error attribution on awareness of the perturbation, using a process dissociation procedure (PDP; Werner et al., 2015; Modchalingam et al., 2019). First, we implemented a baseline-correction procedure, by subtracting no-cursor reaches in the aligned session from no-cursor with strategy and no-cursor without strategy trials, respectively. We then compared angular reach deviations per participant when either including any strategy they learned during adaptation training (no-cursor with strategy) or excluding such a strategy (no-cursor without strategy). We expect that our manipulations will increase participants’ awareness of the external nature of the perturbation and lead to a difference between without strategy and with strategy trials. This is because with strategy trials capture the explicitly available strategy. On the other hand, a lack of awareness will not lead to a difference between without and with strategy trials. We performed a 2X4 mixed design ANOVA with strategy use as a within-subject factor (with or without strategy) and group as a between-subject factor.

```{r}
RAEANOVA()
```

We see that the main effects of group and strategy, as well as their interaction were significant. This suggests that angular reach deviations from at least one group differed from the other groups, when either using or not using a strategy. Planned follow-up tests were necessary to determine which of the groups differed.

```{r}
RAEComparisonMeans()
getRAEComparisonEffSize(method='sidak')
```

The tests confirmed our conclusions from the figure. All groups, except the Control group, showed significantly different angular reach deviations when comparing reaches using a strategy to reaches when they are not using a strategy. This suggests unawareness of learning for the Control group. Note that effect sizes are provided with eta-squared, which is interpreted as the percent of variance in angular reach deviation that is accounted for by the difference between using and not using a strategy.

# Localization

We also investigated how manipulating the extent of external error attribution affected participants' hand location estimates.

we analyzed estimates of hand position in Active Localization and Passive Localization trials, before and after rotation training. We calculated the angular difference between a line connecting the home position to the point where the participant’s unseen right hand movement ended, and a line connecting the home position to the participant’s perceived hand location, i.e. their touchscreen response using the left hand. For us to be able to calculate these angular difference measures appropriately, the touchscreen and monitor position must be coordinated. However, during our analyses of the data, we encountered inconsistencies with the location of the touchscreen responses relative to the arc location. This is because the touchscreen position is fixed, but the monitor can be moved to different positions. Thus, these inconsistencies showed that some of the responses were not centred on where the arc stimuli were expected to be displayed (12 cm from home position). Despite this, responses still fell within the shape of an arc, which allowed us to fix for this problem by implementing a circle fitting procedure that finds a circle of best fit for the responses. Moreover, this procedure also allowed us to account for any participant response biases. The procedure used a fitting algorithm that minimized mean squared errors (MSE), where errors were defined as the difference between the touchscreen response locations and the expected arc stimulus location. The MSE measures were then passed on to an optimizing function (optim in R), which minimized the error to return a best fitting home position, such that the responses were shifted to where we expected it to be (i.e. 12 cm from home position). In short, the circle fitting procedure ensured that any localization shifts we detected in our analyses were not due to technical issues or response biases. Furthermore, we conducted an outlier removal procedure for this fitted data, such that trials with hand
movement endpoints beyond ±20° from the arc centre and angular errors beyond ±3 standard deviations from the mean angular error were removed. As participants chose the points on the arc that they moved towards, their movements did not end on all possible points on the arc. To account for this, we used a kernel smoothing method (gaussian kernel with width = 15°) to interpolate changes in hand localization at specific points
(50°, 90°, and 130°) for every participant. These locations are the same points where the arc stimuli are centred on in polar coordinates. We then used the mean of these interpolated values at each of the three points, to estimate hand localization errors in active and passive movements for both the aligned and rotated sessions.

## Show the data

Below is a plot of the mean angular shift in hand localization for each group as a function of the three interpoints (50°, 90°, and 130°). Panel 1 shows shifts in hand localization when both proprioceptive information and efference-based predictions are available to the participant. Panel 2 shows shifts in hand localization when only proprioceptive information is available. Panel 3 shows shifts in hand localization measured when subtracting passive from active movements. This difference is what we use as a measure for efference-based predictions. Lines represent the mean angular shift in localization for each group, and corresponding shaded regions represent 95% Confidence Intervals. The right hand side of each panel shows the mean localization shifts and corresponding 95% Confidence Intervals for each group, regardless of Hand Angle.

```{r fig.width=8, fig.height=5}
plotLocalizationShift(target='inline')
```

We see from the figure that all groups showed a shift in hand localization, and that these shifts were larger during active localization. In comparing Panels 2 and 3, we see that majority of the shifts in active localization is accounted for with passive localization.

Below is a plot showing individual data points (transparent dots) for shifts in localization according to groups and movement type. For the lower half, solid dots and error bars represent the mean and bootstrapped 95% confidence intervals.

```{r fig.width=8, fig.height=8}
plotLocalizations()
```


## Localization errors differ after rotated training

We first confirmed whether hand position estimates did shift after adaptation to a visuomotor rotation. To do this, we compared
hand localization errors in the aligned session from those in the rotated session. We ran a 2X2X4 mixed design ANOVA with session (*rotated_b* in data; aligned and rotated) and movement type (*passive_b* in data; active and passive) as within-subjects factors and group (instructed, Control, cursor jump, and hand view) as a between-subject factor.

```{r}
localizationANOVA(test='omnibus')
```

We see that rotated_b had a significant main effect, which suggests that hand localization errors after rotated training differed from those before training. We also see other effects which we will investigate further.

## Effect of group and movement type on localization shifts

The other effects show that estimates of hand position do shift despite external error attribution, but the extent of these shifts are modulated by group and movement type. Thus, we ran a 2X4 mixed design ANOVA with movement type (*passive_b*) as a within-subject factor and group as a between-subject factor on localization shifts (i.e. difference in localization error between rotated and aligned sessions). 

```{r}
localizationANOVA(test = 'shifts')
```

We see that there is a main effect of group and movement type, but no significant interaction. This suggests that the effects of movement type on shifts in hand position estimates were not modulated by group, and vice-versa. The effect of movement type is expected, given that active movements have both afferent and efferent-based contributions to hand localization, while passive movements only had afferent information. For the group effect, pairwise comparisons show that the Hand View group differs from the Instructed group.

```{r}
getlocalizationshiftComparisonEffSize()
```


## Proprioceptive Recalibration

We wanted to test whether each group showed shifts in passive localization. We ran t-tests that compared each group's magnitude of shifts in passive localization estimates against 0. These were one-tailed t-tests, where the null hypothesis was that the mean is greater than or equal to 0; the alternative hypothesis was that the mean is less than 0. Hence confidence intervals are all from minus infinity. Note that we do not need to control for multiplicity here, as each group is independently compared to zero (i.e., these are not pairwise comparisons).

```{r}
pasLocTtests()
```

From the tests, we see that all groups differed significantly from 0, suggesting that all groups showed changes in proprioceptive estimates. Note that effect sizes are provided with Cohen's d, which is in standard deviation units. Thus, it indicates how many standard deviations the mean shift for a group is different from zero. Here, all Cohen's d values are relatively similar, except for the Hand View group.


## Predicted Sensory Consequences

Next, we ran t-tests that compared each group's magnitude of changes in predicted sensory consequences against 0.

```{r}
predConsTtests()
```

From the tests, we see that only the Hand View group did not significantly differ from 0, suggesting that the Hand View group had a reduction of updates in predicted sensory consequences. Note that effect sizes are provided with Cohen's d, which is in standard deviation units. Thus, it indicates how many standard deviations the mean shift for a group is different from zero. Here, only the Hand View group has a very small effect size.


# Contributions of Proprioception and Predicted Sensory Consequences on Reach Aftereffects

Next, we tested whether both afferent and efferent-based changes predict/contribute to implicit motor changes. 

To visualize the data, we plot the relationship of passive localization and aftereffects, as well as predictions and aftereffects, individually (a and b in figure). Note that we did perform a Pearson product moment correlation test for these two and found a significant relationship for both, but we only include it as analyses extra to the manuscript (i.e., the main test is the multiple regression below). For the extra analyses, refer to the other R Markdown document, AnalysesSupp.

```{r fig.width=12, fig.height=5}
plotPropPredRelationships()
```

From panels a and b of the plot, we see that there is a relationship between predictions and aftereffects, as well as passive localization and aftereffects. For these plots and the regression below, we pooled data from all individuals across all groups together. One concern could be that we are only detecting a relationship because of a group effect that is leading to a spurious correlation. However, even after taking into account the group effect, these predictors are still statistically significant. (See AnalysesSupp for these tests).

We then performed a multiple regression, where measures of aftereffects was the dependent variable, and both passive localization (or recalibrated proprioception) and predicted sensory consequences were the predictors.

```{r}
getPropPredGLM()
```

We found that both changes in proprioception (Beta = -0.69, p < .001) and predictions (Beta = -0.43, p < .001) were significant predictors of reach aftereffects. As a measure of effect size, we can use the squared semi-partial correlation coefficient (sr-squared). The square root values of sr-squared is just the correlation of the predictor with RAE, after partialling out the effects of the other predictor. Sr-squared, therefore explains the change in total R-squared, when adding in this predictor. This means that 17% of the variance in reach aftereffects is uniquely accounted for by updates in predictions, after partialling out proprioceptive changes. On the other hand, 20% of the variance in reach aftereffects is accounted for by proprioceptive changes, after partialling out updates in predictions. (See AnalysesSupp.Rmd for a discussion of how these are shown with the use of regression models and residuals).

Importantly, we note that changes in proprioception are independent from predicted sensory consequences due to how we have measured them. Recall that proprioceptive recalibration was measured from Passive localization, while predictions were calculated as a difference of Active and Passive Localization. This process ensures predictions do not have any shared variance with proprioceptive recalibration. Although the two components of hand localization are correlated with each other (see correlation test below), predictions and proprioceptive changes are not collinear with each other (vif = 1.09 is considered low; see AnalysesSupp for extra analyses).

```{r}
getPropPredCorrelation()
```


Next, we calculated the predicted reach aftereffects from our model and compared these with the actual/observed reach aftereffects from our experiment (panel c in the plot above).To calculate the predicted reach aftereffects, we refer to values from our model. The equation is:

predicted RAE = Intercept + Prediction_Coeff(actual_prediction) + Proprioception_Coeff(actual_proprioception)

If the model predicted reach aftereffects perfectly, then points would fall along the diagonal of the plot. We see that the predictions overshoot the data on the lower end, but undershoot the data on the higher end. These could possibly be due to noise, or it could mean that the model is incomplete. It is very likely that another predictor needs to be included in the model, but the current study did not look further into this. We note that these results do not definitively show that changes in proprioception and predictions are what cause aftereffects. That is, we cannot establish causation from this multiple regression. If we considered the correlations of each predictor with aftereffects (r-squared values from panels a and b of plot), these values are still small in magnitude. It is possible that the significant relationships we found were due to an alternative variable or a common source (e.g. factors such as participant motivation). Therefore, although our results suggest that both proprioception and predictions independently contribute to aftereffects, these should be interpreted with caution.

Alternatively, we provide further support for the relationships of each component of hand localization with implicit aftereffects, by showing the lack of such a relationship with explicit learning (i.e. with minus without strategy no-cursor trials; panels d to f).

```{r}
getPropPredExpRelationships()
```

Individually, each component does not have a statistically significant relationship with explicit learning. Combining the two independent variables in a multiple regression where explicit learning is the dependent variable, shows that updates in predictions are statistically significant. However, the coefficient indicates that when the efferent component of hand localization goes up, explicit reach aftereffects go down. Regardless, the adjusted r-squared of this multiple regression model is still low. Moreover, we validate the regression model by comapring predicted with actual explicit learning (panel f in plot above). In short, these suggest that proprioceptive recalibration or updates in predicted sensory consequences do not contribute to explicit learning, but seem to contribute to implicit learning.

## Extra analyses for residual structure in regression

we decided to see if transforming the slightly curvilinear data (in passive localization shifts - panel a in plot) could reduce the residuals. We used a square-root transformation for passive localization shifts and re-ran the multiple regression. Doing so increased the adjusted r-squared from 0.276 to 0.365, which we interpret as a significant improvement.

```{r}
getSqrtPasLoc()
```

Even though such a transformation may help to somewhat improve the model, we emphasize that the original (and simpler) model is enough to conclude that proprioceptive recalibration and updated predicted sensory consequences reflect separate processes that are independently related to implicit aftereffects.
